\ttitle

### 2013-11-11 Chapters/Chapter3/one_class_svm.tex
Add UCI HAR plot and other notes

This minimization problem can be solved (using \gls{qp}) and transformed to its (Lagrange) dual formulation.
In the dual formulation the problem scales with the number of training examples $n$ instead of the dimensionality $d$ of the samples.
Solving this problem directly in the high dimensional feature space $\mathcal{F}$ makes it untractable.
The linear approximation function corresponds to the kernel function in the dual formulation.
Solving this dual formulation is equivalent to solving the primal formulation \cite{cherkassky2007learning}.
In the dual formulation the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
\begin{equation}\label{eq:svm_lagrange}
  f(x) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i y_i K(x, x_i) + b),
\end{equation}
where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in Section~\ref{subsec:kernel_trick}).
Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
Since it is shown that under certain circumstances \glspl{svm} show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.

>>>>

This minimization problem can be solved (using \gls{qp}) and transformed to its (Lagrange) dual formulation.
In the dual formulation the problem scales with the number of training examples $n$ instead of the dimensionality $d$ of the samples.
Solving this problem directly in the high dimensional feature space $\mathcal{F}$ makes it untractable.
The linear approximation function corresponds to the kernel function in the dual formulation.
Solving this dual formulation is equivalent to solving the primal formulation \cite{cherkassky2007learning}.
In the dual formulation the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
\begin{equation}\label{eq:svm_lagrange}
  f(x) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i y_i K(x, x_i) + b),
\end{equation}
\TODO{Elaborate on imporant properties of page 56, in Appendix A: maximum-margin is searching for support vectors and pairwise dot products, Kernel matrix.}
where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in Section~\ref{subsec:kernel_trick}).
Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
Since it is shown that under certain circumstances \glspl{svm} show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.

### 2013-11-11 Chapters/Chapter4/change_detection.tex
Add UCI HAR plot and other notes

\subsection{Ratio-thresholding}
The second example of an algorithm that interprets the change indication time series and transforms it to change detection is the radius ratio thresholding method used by Camci~\cite{camci2010change}.
The method shares characteristics with \gls{cusum}, since it relies on summation of historic data and thresholds for change detection.
The method calculates the radius ratio $\hbar_t$ by taking the ratio of the current radius $r_t$ at time $t$ to the average radii since the last change point $y$:
\begin{equation}\label{eq:ratio_radius}
  \hbar_t = \frac{r_t}{\operatorname*{mean}(r_{y:t-1})},
\end{equation}
where $\operatorname*{mean}(r_{y:t-1})$ is the average of the previous approximated radii.
By using only historic data, this method can be incorporated in an online change detection method.
The ratio $\hbar_t$ is compared with two thresholds: the low threshold $th_\text{low}$ and the high threshold $th_\text{high}$.
When the data objects in the working set becomes more homogeneous at time $t$, the radius $R$ of the hypersphere and thereby the ratio $\hbar_t$ will decrease.
When the new value of $\hbar_r$ is lower than the low threshold $th_\text{low}$, a change detection is observed for time $t$.
The same holds for a more heterogeneous set of data and an increase of $\hbar_r$ higher than $th_\text{high}$.

@@

>>>>

\TODO{This subsection has double information. Cleanup. Mention iterated cumulative sum of squares of inclan \cite{inclan1994use}.}

\subsection{Ratio-thresholding}\label{subsec:ratio_thresholding}
The second example of an algorithm that interprets the change indication time series and transforms it to change detection is the radius ratio thresholding method used by Camci~\cite{camci2010change}.
The method shares characteristics with \gls{cusum}, since it relies on summation of historic data and thresholds for change detection.
The method calculates the radius ratio $\hbar_t$ by taking the ratio of the current radius $r_t$ at time $t$ to the average radii since the last change point $y$:
\begin{equation}\label{eq:ratio_radius}
  \hbar_t = \frac{r_t}{\operatorname*{mean}(r_{y:t-1})},
\end{equation}
where $\operatorname*{mean}(r_{y:t-1})$ is the average of the previous approximated radii.
By using only historic data, this method can be incorporated in an online change detection method.
The ratio $\hbar_t$ is compared with two thresholds: the low threshold $th_\text{low}$ and the high threshold $th_\text{high}$.
When the data objects in the working set becomes more homogeneous at time $t$, the radius $R$ of the hypersphere and thereby the ratio $\hbar_t$ will decrease.
When the new value of $\hbar_r$ is lower than the low threshold $th_\text{low}$, a change detection is observed for time $t$.
The same holds for a more heterogeneous set of data and an increase of $\hbar_r$ higher than $th_\text{high}$.

@@

\begin{center}\begin{table}
\begin{tabular}{ l p{12cm} }
  \hline
  Step & Action \\
  \hline
  1 & Start with $n$ data objects as working set and construct hypersphere \\
  2 & Add next data object $x_t$ and drop first one \\
  3 & Identify new hyper-sphere and its \emph{approximate radius} \\
  4 & calculate radius average of hyperspheres since last change point \\
  5 & calculate radius ratio $\hbar$. If lower than ${th}_{low}$ or greater than ${th}_{high}$ then mark $t$ as change point \\
  6 & continue from step 2, until no more data objects available \\
  7 & Apply post-processing: merge close change points \\
  \hline
\end{tabular}
\caption{Algorithm of the proposed change detection method, using the Ratio-Thresholding method of \ref{subsec:ratio_thresholding} for change detection.}
\label{tab:algorithm_proposed_method}
\end{table}\end{center}

\TODO{Add reference to Table~\ref{tab:algorithm_proposed_method}.}

### 2013-11-11 Chapters/Chapter6/data_gathering.tex
Add UCI HAR plot and other notes

\TODO{Use common data sets. The WISDM dataset is not usefull because it has no transitions, only very clear blocks of the same activities. Include a plot to illustrate the problem?}

\begin{figure}\label{fig:wisdm_excerpt}
\centering
  \includegraphics[width=1\textwidth]{./Figures/Chapter6/data_collection/wisdm_excerpt.eps}
  \caption[Jumping variance ratios]{Excerpt from the WISDM data set \cite{kwapisz2011activity}. First five activities for subject $33$ are displayed. Due to the discontinuous nature of the recordings, this data set can not be used for change detection.}
\end{figure}

>>>>

\TODO{Use common data sets. The WISDM dataset is not useful because it has no transitions, only very clear blocks of the same activities. Include a plot to illustrate the problem?}

\begin{figure}\label{fig:wisdm_excerpt}
\centering
  \includegraphics[width=1\textwidth]{./Figures/Chapter6/data_collection/wisdm_excerpt.eps}
  \caption[WISDM Excerpt]{Excerpt from the WISDM data set \cite{kwapisz2011activity}. First five activities for subject $33$ are displayed. Due to the discontinuous nature of the recordings, this data set can not be used for change detection.}
\end{figure}

\begin{figure}\label{fig:uci_annotated}
\centering
  \includegraphics[width=1\textwidth]{./Figures/Chapter6/data_collection/uci_annotated.eps}
  \caption[UCI HAR Excerpt]{Excerpt from the UCI HAR data set \cite{anguita2012human}. The recorded training activities for subject $25$ are displayed. Due to the imprecise (and possibly incorrect) labeling and the short duration of activities, the data set can not be used for the change detection method as described in this thesis.}
\end{figure}
